---
title: "roahd"
author: "Nicholas Tarabelloni"
date: "`r Sys.Date()`"
output:: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

Package **roahd** (**RO**bust **A**nalysis of **H**igh dimensional **D**ata) is 
an `R` package meant to gather recently proposed statistical methods to deal 
with the robust analysis of functional data.

The package contains an implementation of quantitative methods, based on
functional depths and indices, on top of which are built some graphical methods
(funcitonal boxplot and outliergram) useful to carry out an explorative analysis
of a functional dataset and to robustify it from shape and magnitude outliers.

Both univariate and multivariate functional data are supported in the package,
whose functions have been implemented with a particular emphasis on computational
efficiency, in order to allow the processing of high-dimensional dataset.

## Representation of functional data

The package has been designed to work with a representation of functional data
through dedicated, simple and handy `S3` classes, namely `fData` for univariate
functional data and `mfData` for multivariate functional data.

The use of `S3` classes is exploited by suitable `S3` methods implementing the
statistical features of the package, that are able to dispatch the correct
method call depending on the class of their functional data argument.

### fData

`S3` class `fData` implements a representaion of univariate functional datasets.
They are obtained once specifying, for each observation in the functional
dataset, a set of measurements over a discrete grid, representing the continuous
variable indexing the functional dataset (e.g. time).

In other words, if we denote by $T = [t_0, t_1, \ldots, t_{P-1}]$ an evenly spaced
grid ($t_i - t_{i-1} = h > 0$), and imagine to deal with a dataset 
$D_{i,j} = X_i(t_j)$, $\forall i = 1, \ldots, N$ and $\forall j=0, \ldots, P-1$,
the object `fData` is built starting from the grid and the values in the 
following way: 

```{r fData, fig.align='center', fig.height=4, fig.width=4, cache=TRUE}

library( roahd )

# The number of observations in the functional dataset.
N = 5
# The number of points in the 1D grid where the functional data are measured.
P = 1e2
# The previous two variable names are used consistently throughout the tutorial
# and the package's documentation to indicate the number of observations and the
# grid size.

# The grid over which the functional dataset is defined
grid = seq( 0, 1, length.out = P )

# Creating the values of the functional dataset
Data = matrix( c( sin( 2 * pi * grid  ),
                  cos( 2 * pi * grid ),
                  4 * grid * ( 1 - grid ),
                  tan( grid ),
                  log( grid ) ),
              nrow = N, ncol = P, byrow = TRUE )

# Building an fData object
# The constructor takes a grid and a matrix-like structure for data values
# (see help for more details on how to use the constructor)
fD = fData( grid, Data )

# Inspecting the structure of an fData object
str( fD )

plot( fD, main = 'Univariate FD', xlab = 'time [s]', ylab = 'values', lwd = 2 )
```
Thus, an object `fData` is a list containing: the fields `t0`, `tP`, the defining
the starting and end point of the one dimensional grid of the `fData`, the
constant step size `h` and the number of grid points `P`; the field 

### mfData

An `mfData` object, instead, can be created in this way:

```{r mfData, cache = TRUE, fig.width = 7, fig.height = 4, fig.align = 'center' }

# Creating some values for first component of the dataset
Data_1 = t( sapply( runif( 10, 0, 4 ), 
                    function( phase ) sin( 2 * pi * grid + phase ) ) )

# Creating some values of functions for  
Data_2 = t( sapply( runif( 10, 0, 4 ), 
                    function( phase ) log( grid + phase ) ) )


# Building an fData object
# The constructor takes a grid and a list of matrix-like structures for data values
# (see help for more details on how to use the constructor)
mfD = mfData( grid, list( Data_1, Data_2 ) )

str( mfD )

# Each component of the mfData object is a fData object 
sapply( mfD$fDList, class )


plot( mfD, lwd = 2, main = 'Multivariate FD',
      xlab = 'time', ylab = list( 'Values 1', 'Values 2' ))

```

The fact that `mfData` components are `fData` obejcts is indeed conceptyally
very natural, but also allows for a seamless application of `S3` methods meant 
for `fData` on multivariate functional data components, making the exploration
and manipulation of multivariate datasets rather easy:

```{r mfDataComponents, fig.width = 4, fig.height = 4, fig.align = 'center' }

plot( mfD$fDList[[ 1 ]], main = 'First component', 
      xlab = 'time', ylab = 'Values', lwd = 2 )

# S3 method for mean of univariate functional data
plot( mean( mfD$fDList[[ 1 ]] ), lwd = 2, col = 'black', add = TRUE )
```

Moreover, `mfData` objects can be obtained also from a set of homogeneous 
`fData` objects, i.e. of equal sample size and defined on the same grid:

```{r fDataToMfData, eval = FALSE }

fD_1 = fData( grid, Data_1 )

fD_2 = fData( grid, Data_2 )

mfD = as.mfData( list( fD_1, fD_2 ) ) 

mfD = as.mfData( lapply( 1 : 10, function( i )( fD_1 ) ) )
```

### Subsetting fData

`fData` objects can be subset using a suitably overloaded operator `[.fData`,
that allows for the use of standard slices of `matrix` and `array` classes also
for `fData`.

```{r fData_subset_1, eval = FALSE }

# Subseting fData and returning result in matrix form
fD[ 1 , 1, as_fData = FALSE ]

fD[ 1, , as_fData = FALSE ]

fD[ 2, 10 : 20, as_fData = FALSE ]

fD[ , 10, as_fData = FALSE ]
```
```{r fData_subset_2, fig.align='center', fig.width=7, fig.height=4, cache=TRUE, }
# As default behaviour the subset is returned in fData form
par( mfrow = c(1,2) )
plot( fD, main = 'Original dataset', lwd = 2 )
plot( fD[ , 1 : 20 ], main = 'Zooming in', lwd = 2 )
```

### Algebra

An algebra of `fData` obejcts is also implemented, making it easy to sum,
subtract, multiply and divide these objects by meaningful and compliant
structures.

**Sums** and **subtractions**, available through `+` and `-` operators (see help at
`+-.fData`), allow for sums of a `fData` on the left hand side and a compliant
structure on the right hand side. This can be either another `fData` of same
sample size and defined over the same grid, or a 1D/2D data structure with a
number of columns equal `fData`'s grid length (i.e. `P`), and number of rows
equal to `fData`'s sample size (i.e. `N`) or equal to one (in this case the only
observation available is recycled `N` times). The operataions are then performed
element-wise between the lhs and rhs. 

```{r fData_algebra_1, eval = FALSE }
fD + fD

fD + matrix( 1, nrow = N, ncol = P )

fD + array( 2, dim = c( N, P ) )

fD + 1 : P 
```
**Multiplication** and **division**, instead, is implemented only for an `fData` 
left hand side and a numeric variable or numeric vector right hand side. In the 
first case, each function in the functional dataset is multiplied/divided by the
specified quantity; in the second case, specifying a vector of length `N`, the
multiplication/division of each functional observation is carried out by the
corresponding quantity in the vector, in an element-wise way:

```{r fData_algebra_2, eval = FALSE}
fD * 2

fD / 3

fD * ( 1 : N )

fD / ( 1 : N )
```

## Statistics

### Depths
`roahd` provides a number of depth definitions, that are exploited by the
visualization functions but can also be used by themselves. These are based on
the notion of **Band Depth** [@lopezpintado_romo_2007; @lopezpintado_romo_2009].

**Band Depth** and **Modified Band Depth** are implemented in the functions
`BD` and `MBD`. Both of them can work with either an `fData` object, specifying
the univariate functional dataset whose depths must be computed, or a matrix of
data values (e.g. in the form of `fData$values` output). 
`MBD` can be called with the additional parameters `manage_ties` (defaulting to
`FALSE`), specifying whether a check for the presence of tied data must be made
prior to computing depths, and therefore a suitable computing strategy must
be used. 
The implementation of `MBD` exploits the recommendations of [@Sun_Genton_2012],
but extends them in order to accomodate for the possible presence of ties.

```{r BD_MBD_1, eval = FALSE, cache = TRUE }

  BD( fD )
  BD( fD$values )
  
  MBD( fD )
  MBD( fD$values )
  
  MBD( fD, manage_ties = TRUE )
  MBD( fD$values, manage_ties = TRUE )
```

Another definition avaibable is the **Half Region Depth** (along with its modified
version), that is built on top of the **Epigraph** and **Hypograph** indexes [
see @lopezpintado_romo_2011]:

```{r HRD, eval = FALSE, cache = TRUE}

HRD( fD )
HRD( fD$values )

MHRD( fD )
MHRD( fD$values )

```

A generalisation of MBD to multivariate functional data, implementing the ideas
of @ieva_paganoni_2013, is also available thorugh the functions `multiBD` and
`multiMBD`. These functions accept either a `mfData` object, specifying the
multivariate functional dataset whose depths have to be computed, or a list of
matrices of data values (see example below).

The function computes the BD or MBD for each component of the multivariate
functional dataset and then averages them accorging to a set of weights.
These can be specified in two ways: either with the flag `uniform`, that is
turned into `rep( 1/L, L )`, or by providing the actual set of weights to be
used. The latter option allows to use ad-hoc set of weights, like in 
[@tarabelloni_ieva_etal_2015].

```{r multiMBD, eval = FALSE, cache = TRUE }

  multiBD( mfD, weights = 'uniform' )
  multiMBD( mfD, weights = 'uniform', manage_ties = FALSE )

  multiBD( mfD, weights = c( 0.6, 0.4) )
  multiMBD( mfD, weights = c( 0.7, 0.3 ), manage_ties = FALSE )

  multiBD( list( fD_1$values, fD_2$values ), weights = c( 0.6, 0.4) )
  multiMBD( list( fD_1$values, fD_2$values ), weights = c( 0.7, 0.3 ), manage_ties = FALSE )
```

### Mean and median
Suitable `S3` extensions to the functions for the computation of mean function 
and median of functional datasets are also present.

The sample mean of a functional dataset coincides with the __cross-sectional__
mean function, i.e. the function obtained by computing the mean across the whole
dataset point-by-point along the grid where functional data are defined:

$$
\widehat{\mu}( t_j ) = \dfrac{1}{N} \sum_{i=1}^{N} X_i(t_j), \qquad \forall j = 0, \ldots, P-1
$$

for a univariate functional dataset, while in the multivariate case is:

$$
  \widehat{\mu}_k(t_j) = \dfrac{1}{N} \sum_{i=1}^{N} X_{i,k}(t_j), \qquad \forall j = 0, \ldots, P-1,
  \qquad \forall k = 1, \ldots, L.
$$

The sample median of a functional dataset, instead, is defined as the element of
the functional dataset fulfilling the maximum depths, given a certain definition
of depth. For instance, for MBD:

$$
  \widehat{m}( t_j ) = \left(\text{arg}\max_{i=1, \ldots, N} MBD( X_i )\right)( t_j), \qquad j = 0, \ldots, P-1.
$$

The sample mean is implemented in the `S3` methods `mean.fData( x, ... )` and
`mean.mfData( x, ... )`, that can be invoked directly on `fData` and `mfData`
objects.
The sample median, instead, is implemented in the functions 
`mean_fData( fData, type = 'MBD', ... )` and 
`mean_mfData( mfData, type = 'multiMBD', ... )`. Here, the `type` flag can take
any name of functions (available in the caller's environment) that can be used to
compute the depths defining the sample median, taking respectively a 2D matrix of
data values and a list of 2D matrix of data values as argument (plus, optionally,
the dots arguments).

```{r, eval = FALSE, cache = TRUE }

  mean( fD )
  mean( mfD)
  
  median_fData( fD, type = 'MBD' )
  median_fData( fD, type = 'MHRD' )

```
```{r, eval = TRUE, fig.height=3, fig.width=6, fig.align='center'}

  par( mfrow = c(1,2) )
  plot( fD, main = 'Mean', lwd = 2 )
  plot( mean( fD ), add = TRUE, lwd = 2, col = 'black', lty = 2 )
  
  plot( fD, main = 'Median', lwd = 2 )
  plot( median_fData( fD, type = 'MBD' ), add = TRUE, lwd = 2, lty = 2, 
        col = 'black' )
```





### Indexes


### Correlation coefficients


## Simulation of functional data

## Visualization




## References
