---
title: "roahd"
author: "Nicholas Tarabelloni"
date: "`r Sys.Date()`"
output:: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

Package **roahd** (**RO**bust **A**nalysis of **H**igh dimensional **D**ata) is 
an `R` package meant to gather recently proposed statistical methods to deal 
with the robust analysis of functional data.

The package contains an implementation of quantitative methods, based on
functional depths and indices, on top of which are built some graphical methods
(funcitonal boxplot and outliergram) useful to carry out an explorative analysis
of a functional dataset and to robustify it from shape and magnitude outliers.

Both univariate and multivariate functional data are supported in the package,
whose functions have been implemented with a particular emphasis on computational
efficiency, in order to allow the processing of high-dimensional dataset.

## Representation of functional data

The package has been designed to work with a representation of functional data
through dedicated, simple and handy `S3` classes, namely `fData` for univariate
functional data and `mfData` for multivariate functional data.

The use of `S3` classes is exploited by suitable `S3` methods implementing the
statistical features of the package, that are able to dispatch the correct
method call depending on the class of their functional data argument.

### fData

`S3` class `fData` implements a representaion of univariate functional datasets.
They are obtained once specifying, for each observation in the functional
dataset, a set of measurements over a discrete grid, representing the continuous
variable indexing the functional dataset (e.g. time).

In other words, if we denote by $T = [t_0, t_1, \ldots, t_{P-1}]$ an evenly spaced
grid ($t_i - t_{i-1} = h > 0$), and imagine to deal with a dataset 
$D_{i,j} = X_i(t_j)$, $\forall i = 1, \ldots, N$ and $\forall j=0, \ldots, P-1$,
the object `fData` is built starting from the grid and the values in the 
following way: 

```{r fData, fig.align='center', fig.height=4, fig.width=4, cache=TRUE}

library( roahd )

# The number of observations in the functional dataset.
N = 5
# The number of points in the 1D grid where the functional data are measured.
P = 1e2
# The previous two variable names are used consistently throughout the tutorial
# and the package's documentation to indicate the number of observations and the
# grid size.

# The grid over which the functional dataset is defined
grid = seq( 0, 1, length.out = P )

# Creating the values of the functional dataset
Data = matrix( c( sin( 2 * pi * grid  ),
                  cos( 2 * pi * grid ),
                  4 * grid * ( 1 - grid ),
                  tan( grid ),
                  log( grid ) ),
              nrow = N, ncol = P, byrow = TRUE )

# Building an fData object
# The constructor takes a grid and a matrix-like structure for data values
# (see help for more details on how to use the constructor)
fD = fData( grid, Data )

# Inspecting the structure of an fData object
str( fD )

plot( fD, main = 'Univariate FD', xlab = 'time [s]', ylab = 'values', lwd = 2 )
```
Thus, an object `fData` is a list containing: the fields `t0`, `tP`, the defining
the starting and end point of the one dimensional grid of the `fData`, the
constant step size `h` and the number of grid points `P`; the field 

### mfData

An `mfData` object, instead, can be created in this way:

```{r mfData, cache = TRUE, fig.width = 7, fig.height = 4, fig.align = 'center' }

# Creating some values for first component of the dataset
Data_1 = t( sapply( runif( 10, 0, 4 ), 
                    function( phase ) sin( 2 * pi * grid + phase ) ) )

# Creating some values of functions for  
Data_2 = t( sapply( runif( 10, 0, 4 ), 
                    function( phase ) log( grid + phase ) ) )


# Building an fData object
# The constructor takes a grid and a list of matrix-like structures for data values
# (see help for more details on how to use the constructor)
mfD = mfData( grid, list( Data_1, Data_2 ) )

str( mfD )

# Each component of the mfData object is a fData object 
sapply( mfD$fDList, class )


plot( mfD, lwd = 2, main = 'Multivariate FD',
      xlab = 'time', ylab = list( 'Values 1', 'Values 2' ))

```

The fact that `mfData` components are `fData` obejcts is indeed conceptyally
very natural, but also allows for a seamless application of `S3` methods meant 
for `fData` on multivariate functional data components, making the exploration
and manipulation of multivariate datasets rather easy:

```{r mfDataComponents, fig.width = 4, fig.height = 4, fig.align = 'center' }

plot( mfD$fDList[[ 1 ]], main = 'First component', 
      xlab = 'time', ylab = 'Values', lwd = 2 )

# S3 method for mean of univariate functional data
plot( mean( mfD$fDList[[ 1 ]] ), lwd = 2, col = 'black', add = TRUE )
```

Moreover, `mfData` objects can be obtained also from a set of homogeneous 
`fData` objects, i.e. of equal sample size and defined on the same grid:

```{r fDataToMfData, eval = FALSE }

fD_1 = fData( grid, Data_1 )

fD_2 = fData( grid, Data_2 )

mfD = as.mfData( list( fD_1, fD_2 ) ) 

mfD = as.mfData( lapply( 1 : 10, function( i )( fD_1 ) ) )
```

### Subsetting fData

`fData` objects can be subset using a suitably overloaded operator `[.fData`,
that allows for the use of standard slices of `matrix` and `array` classes also
for `fData`.

```{r fData_subset_1, eval = FALSE }

# Subseting fData and returning result in matrix form
fD[ 1 , 1, as_fData = FALSE ]

fD[ 1, , as_fData = FALSE ]

fD[ 2, 10 : 20, as_fData = FALSE ]

fD[ , 10, as_fData = FALSE ]
```
```{r fData_subset_2, fig.align='center', fig.width=7, fig.height=4, cache=TRUE, }
# As default behaviour the subset is returned in fData form
par( mfrow = c(1,2) )
plot( fD, main = 'Original dataset', lwd = 2 )
plot( fD[ , 1 : 20 ], main = 'Zooming in', lwd = 2 )
```

### Algebra

An algebra of `fData` obejcts is also implemented, making it easy to sum,
subtract, multiply and divide these objects by meaningful and compliant
structures.

**Sums** and **subtractions**, available through `+` and `-` operators (see help at
`+-.fData`), allow for sums of a `fData` on the left hand side and a compliant
structure on the right hand side. This can be either another `fData` of same
sample size and defined over the same grid, or a 1D/2D data structure with a
number of columns equal `fData`'s grid length (i.e. `P`), and number of rows
equal to `fData`'s sample size (i.e. `N`) or equal to one (in this case the only
observation available is recycled `N` times). The operataions are then performed
element-wise between the lhs and rhs. 

```{r fData_algebra_1, eval = FALSE }
fD + fD

fD + matrix( 1, nrow = N, ncol = P )

fD + array( 2, dim = c( N, P ) )

fD + 1 : P 
```
**Multiplication** and **division**, instead, is implemented only for an `fData` 
left hand side and a numeric variable or numeric vector right hand side. In the 
first case, each function in the functional dataset is multiplied/divided by the
specified quantity; in the second case, specifying a vector of length `N`, the
multiplication/division of each functional observation is carried out by the
corresponding quantity in the vector, in an element-wise way:

```{r fData_algebra_2, eval = FALSE}
fD * 2

fD / 3

fD * ( 1 : N )

fD / ( 1 : N )
```

### Visualization
`fData` and `mfData` objects can be visualised thanks to specific `S3` plotting
methods, `plot.fData` and `plot.mfData`.

The graphical parameters of these functions have been suitably customised in order
to enhance the visualization of functions. In particular, as a default elements
are plotted with continuous lines and an ad-hoc palette that helps differentiating
them. As default x- and y-axis labels, as long as titles, are dropped so that
plot's arguments calls are not displayed when no value is provided.

In case of `mfData` the graphical window is split into a rectangular lattice to
plot single dimensions. The rectangular frame has `floor( sqrt( L ) )` rows and
`ceiling( x$L / floor( sqrt( x$L ) ) )` columns. If custom labels/titles are
desired, they must be provided in the following way: since the grid is the same
for all the dimensions, just one string is expected for x-axis
(e.g. `xlab = 'grid'`), while either a single string or a list of `L` strings 
(one for each dimension) is expected for both the y-label(s) and title(s). 
In case just one string is passed to `plot.mfData`, the same value is used for 
all the dimensions.

## Statistics

### Depths
`roahd` provides a number of depth definitions, that are exploited by the
visualization functions but can also be used by themselves. These are based on
the notion of **Band Depth** [@lopezpintado_romo_2007; @lopezpintado_romo_2009].

**Band Depth** and **Modified Band Depth** are implemented in the functions
`BD` and `MBD`. Both of them can work with either an `fData` object, specifying
the univariate functional dataset whose depths must be computed, or a matrix of
data values (e.g. in the form of `fData$values` output). 
`MBD` can be called with the additional parameters `manage_ties` (defaulting to
`FALSE`), specifying whether a check for the presence of tied data must be made
prior to computing depths, and therefore a suitable computing strategy must
be used. 
The implementation of `MBD` exploits the recommendations of [@Sun_Genton_2012],
but extends them in order to accomodate for the possible presence of ties.

```{r BD_MBD_1, eval = FALSE, cache = TRUE }

  BD( fD )
  BD( fD$values )
  
  MBD( fD )
  MBD( fD$values )
  
  MBD( fD, manage_ties = TRUE )
  MBD( fD$values, manage_ties = TRUE )
```

Another definition avaibable is the **Half Region Depth** (along with its modified
version), that is built on top of the **Epigraph** and **Hypograph** indexes [
see @lopezpintado_romo_2011]:

```{r HRD, eval = FALSE, cache = TRUE}

HRD( fD )
HRD( fD$values )

MHRD( fD )
MHRD( fD$values )

```

A generalisation of MBD to multivariate functional data, implementing the ideas
of @ieva_paganoni_2013, is also available thorugh the functions `multiBD` and
`multiMBD`. These functions accept either a `mfData` object, specifying the
multivariate functional dataset whose depths have to be computed, or a list of
matrices of data values (see example below).

The function computes the BD or MBD for each component of the multivariate
functional dataset and then averages them accorging to a set of weights.
These can be specified in two ways: either with the flag `uniform`, that is
turned into `rep( 1/L, L )`, or by providing the actual set of weights to be
used. The latter option allows to use ad-hoc set of weights, like in 
[@tarabelloni_ieva_etal_2015].

```{r multiMBD, eval = FALSE, cache = TRUE }

  multiBD( mfD, weights = 'uniform' )
  multiMBD( mfD, weights = 'uniform', manage_ties = FALSE )

  multiBD( mfD, weights = c( 0.6, 0.4) )
  multiMBD( mfD, weights = c( 0.7, 0.3 ), manage_ties = FALSE )

  multiBD( list( fD_1$values, fD_2$values ), weights = c( 0.6, 0.4) )
  multiMBD( list( fD_1$values, fD_2$values ), weights = c( 0.7, 0.3 ), manage_ties = FALSE )
```

### Mean and median
Suitable `S3` extensions to the functions for the computation of mean function 
and median of functional datasets are also present.

The sample mean of a functional dataset coincides with the *cross-sectional*
mean function, i.e. the function obtained by computing the mean across the whole
dataset point-by-point along the grid where functional data are defined:

$$
\widehat{\mu}( t_j ) = \dfrac{1}{N} \sum_{i=1}^{N} X_i(t_j), \qquad \forall j = 0, \ldots, P-1
$$

for a univariate functional dataset, while in the multivariate case is:

$$
  \widehat{\mu}_k(t_j) = \dfrac{1}{N} \sum_{i=1}^{N} X_{i,k}(t_j), \qquad \forall j = 0, \ldots, P-1,
  \qquad \forall k = 1, \ldots, L.
$$

The sample median of a functional dataset, instead, is defined as the element of
the functional dataset fulfilling the maximum depths, given a certain definition
of depth. For instance, for MBD:

$$
  \widehat{m}( t_j ) = \left(\text{arg}\max_{i=1, \ldots, N} MBD( X_i )\right)( t_j), \qquad j = 0, \ldots, P-1.
$$

The sample mean is implemented in the `S3` methods `mean.fData( x, ... )` and
`mean.mfData( x, ... )`, that can be invoked directly on `fData` and `mfData`
objects.
The sample median, instead, is implemented in the functions 
`mean_fData( fData, type = 'MBD', ... )` and 
`mean_mfData( mfData, type = 'multiMBD', ... )`. Here, the `type` flag can take
any name of functions (available in the caller's environment) that can be used to
compute the depths defining the sample median, taking respectively a 2D matrix of
data values and a list of 2D matrix of data values as argument (plus, optionally,
the dots arguments).

```{r mean_1, eval = FALSE, cache = TRUE }

  mean( fD )
  mean( mfD)
  
  median_fData( fD, type = 'MBD' )
  median_fData( fD, type = 'MHRD' )

```
```{r mean_2, eval = TRUE, fig.height=3, fig.width=6, fig.align='center'}

  par( mfrow = c(1,2) )
  plot( fD, main = 'Mean', lwd = 2 )
  plot( mean( fD ), add = TRUE, lwd = 2, col = 'black', lty = 2 )
  
  plot( fD, main = 'Median', lwd = 2 )
  plot( median_fData( fD, type = 'MBD' ), add = TRUE, lwd = 2, lty = 2, col = 'black' )
```

### Indexes
`roahd` collects the implementation of some useful indexes that can be used to 
describe and summarise functional datasets. 

`EI` and `MEI` implement the **Epigraph Index** and the **Modified Epigraph Index**,
while `HI` and `MHI` implement the **Hypograph Index** and the **Modified Hypograph Index**
[see @lopezpintado_romo_2011; @arribasgil_romo_2014].
These indexes can be used to sort data, respectively, in a top-down and bototm-up
fashion, and are used to define the HRD/MHRD and to build the outliegram.

These `S3` methods can be called on univariate functional datasets, provided
either in form of a `fData` object or a 2D matrix of values.

```{r indexes, eval = FALSE }
  
  # Calling on fData objects
  EI( fD )
  MEI( fD )
  
  HI( fD )
  MHI( fD )

  # Calling on 2D matrix type objects
  EI( fD$values )
  EI( matrix( rnorm( 20 ), nrow = 4, ncol = 5 ) )
  
```

### Correlation coefficients

When dealing with multivariate functional data, in particular in case of **bivariate**
data, it is possible to compute correlation coefficients between data components
that generalise the Kendall's tau and Spearman's coefficients **REFERENCES!**.

The function `cor_kendall( mfD, ordering = 'max' )` allows to compute the 
Kendall's tau correlation coefficient between components of a bivariate dataset.
The function accepts a `mfData` object and a criterion to perform the ordering
of functional data (this ordering is used to determine the concordances and
discordacnes between pairs in the definition of the coefficient).

Two criteria are available so far, that directly reflect those proposed in the
reference paper: `max`, for the ordering between maxima of functions,
and `area`, for the ordering between area-under-the-curve of functions.

```{r cor_kendall_1, eval = TRUE, cache = TRUE}
  N = 10
  P = 1e3

  grid = seq( 0, 1, length.out = P )
  
  Data_1 = t( sapply( 1 : N, function( i )( sin( 2 * pi * grid ) + i ) ) )
  # Monotone nonlinear transformation of data
  Data_2 = Data_1^3
  
  mfD = mfData( grid, list( Data_1, Data_2 ) )
```

```{r cor_kendall_2, eval = TRUE, cache = TRUE, collapse = TRUE, fig.align = 'center', fig.height = 4, fig.width = 6} 
  plot( mfD, main = list( 'Comp. 1', 'Comp. 2') )  

  # Kendall correlation of monotonically dependent data is exactly 1
  cor_kendall( mfD, ordering = 'max' )
  cor_kendall( mfD, ordering = 'area' )
```

The funciton `cor_spearman( mfD, ordering = 'MEI', ... )` can be used to compute
the Spearman correlation coefficient for a bivariate `mfData` object, uning the
ordering policy specified by `ordering` (defaulting to `MEI`) to rank univariate
compoenents and then compute the correlation coefficient. Besides `MEI`, also `MHI`
can be used to determine ranks of univariate components. Additional parameters
captured by `...` are passed directly to the rank function. 

```{r Spearman, eval = TRUE, cache = TRUE, collapse = TRUE }
# Spearman correlation of monotonically dependent data is exactly 1
  cor_spearman( mfD, ordering = 'MEI' )
  cor_spearman( mfD, ordering = 'MHI' )
```

## Simulation of functional data
`roahd` contains also some functions that can be used to simulate artificial data
sets of functional data, both univariate and multivariate. These are used in the
adjustment procedure of the outliegram and functional boxplot, but can also be
used to help the development of new methodologies and help their testing.

Artificial univariate data are obtained simulating realisations of a gaussian 
process over a discrete grid with a specific covariance function and center
(e.g. mean or median).
Given a covariance function, $C(s,t)$ and a centerline $m(t)$, the model
generating data is:
$$X_i(t) = m(t) + \epsilon(t), \quad Cov(\epsilon(s),\epsilon(t)) = C(s,t), \quad i = 1, \ldots, N.$$

### Univariate functional data

The function `generate_gauss_fdata(N, centerline, Cov = NULL, CholCov = NULL)` 
can be used to simulate a population of such gaussian functional data. 
The required arguments are: `N` the number of elements to be generated; 
`centerline`, the center of the distribution (mean or median); `Cov`, a matrix
representation of the desired covariance function, meant as the measurements
of such function over a tensor grid whose marginal will be the grid where the
functional data will be defined (i.e. the argument `grid` in `fData`);
`CholCov` the Cholesky factor of the discrete representation of the Covariance
function over the tensor grid (optional and alternative to the argument `Cov`).
The inner procedure to generate the synthetic population of gaussian functional
data makes use of the Cholesky factor of `Cov`, hence by providing its Cholesky
factor, if already present in the caller's scope, can save computing time.

A built-in function can be used to generate exponential-like covariance functions,
namely `exp_cov_function( grid, alpha, beta )`, generating the discretised
version of a covariance of the form $C(s,t) = \alpha e^{-\beta | s - t | }$
over a lattice given by the tensorisation of grid in `grid`.

A comprehensive example is the following:

```{r simulation_univariate, execute = FALSE, collapse = TRUE, fig.align = 'center', fig.width = 7, fig.height = 4}
  N = 50
  P = 1e3

  grid = seq( 0, 1, length.out = P )
  
  Cov = exp_cov_function( grid, alpha = 0.2, beta  = 0.3 )
  
  Data = generate_gauss_fdata( N, centerline = sin( 2 * pi * grid ), Cov = Cov )
  
  fD = fData( grid, Data )
  
  plot( fD, main = 'Gaussian fData', xlab = 'grid', lwd = 2)
```

### Multivariate functional data
The function `generate_gauss_mfdata( N, L, centerline, correlations, listCov = NULL, listCholCov = NULL)` can be used to generate a gaussian dataset of multivariate functional
data. The model generating data is the following:

$$ X_{i,k} = m_k(t) + \epsilon_k(t), \quad Cov(\epsilon_k(s),\epsilon_k(t))=C(s,t), \quad
\forall i = 1, \ldots, N, \quad \forall k = 1, \ldots, L,$$

where $Cor( \epsilon_j(t),\epsilon_l(t))=\rho_{j,l}$ specifies a (*synchronous*) 
correlation structure among the compoenents of the functional dataset.

In order to use the function one should provide: `N`, the number of elelemnts to
simulate; `L`, the number of components of the functional data; `centerline`, a
matrix containing (by rows) the centerline for each component; `correlations`,
a vector of length `1/2 * L * ( L - 1 )`, containing all the correlation 
coefficients among the components; either `listCov` or `listCholCov`, a list
containing either the discretised covariance functions over the tensorised
grid where functional data will be defined), or their Cholesky factor.

A comprehensive example is the following:
```{r simulation_multivariate, execute = FALSE, collapse = TRUE, fig.align = 'center', fig.width = 7, fig.height = 4}
  N = 10
  P = 1e3

  grid = seq( 0, 1, length.out = P )
  
  Cov_1 = exp_cov_function( grid, alpha = 0.1, beta  = 0.5 )
  Cov_2 = exp_cov_function( grid, alpha = 0.5, beta = 0.1)
  
  centerline = matrix( c( sin( 2 * pi * grid ),
                          cos( 2 * pi * grid ) ), nrow = 2, byrow = TRUE )
  
  Data = generate_gauss_mfdata( N, 2, centerline, 0.8, list( Cov, Cov ) )
  
  mfD = mfData( grid, Data )
  
  plot( mfD, main = list( 'Comp.1', 'Comp. 2'), xlab = 'grid', lwd = 2)
```

## Robustification

### Functional boxplot

An implementation of the functional boxplot, through the `S3` method `fbplot`, 
allows for the detection of *amplitude* outliers in univariate and multivariate 
functional datasets [see @Sun_Genton_2011].

`fbplot` can be used to compute the set of indices inside a dataset marking
outlying observations. If used in graphical way (default behaviour), it also 
plots the functional boxplot of the dataset provided.

The functional boxplot is obtained by ranking functions from the center of the
distribution outwards thanks to a depth definition, computing the region of 50%
most central functions and inflating such region by a factor `F`. Whichever
function crossing these boundaries is flagged as outlier. The default value for
`F` is `1.5`, otherwise it can be set with the argument `Fvalue`.

The argument `Depths` can take either the name of the function to call in order
to compute the depths (default is `MBD`), or a vector containing the dephts 
values for the provided dataset.

An example is:

```{r fbplot_fData, fig.align='center', fig.height=5, fig.width=7, cache=TRUE, results='hide'}
  N = 1e2
  P = 1e2
  
  grid = seq( 0, 1, length.out = P )
  
  Cov = exp_cov_function( grid, alpha = 0.2, beta = 0.3 )
  
  Data = generate_gauss_fdata( N, sin( 2 * pi * grid ), Cov )
  
  fD = fData( grid, Data )
  
  Data = generate_gauss_mfdata( N, 2, matrix( sin( 2 * pi * grid ), nrow = 2, ncol = P, byrow = TRUE ), 0.6, listCov = list( Cov, Cov ) )

  mfD = mfData( grid, Data )
  
  fbplot( fD, main = 'Fbplot', Fvalue = 3.5 )

  fbplot( mfD, main = list( 'Comp. 1', 'Comp. 2' ), Fvalue = 3.5 )
```

The method `fboplot.fData` also allows to automatically compute the best 
adjustment factor `F` that yields a desired proportion of outliers 
(True Positive Rate, `TPR`) out of a Gaussian dataset with same center and 
covariance function as the `fData` object [see @Sun_Genton_2012_boxplot].

Such automatic tuning involves the simulation of a number `N_trials` of
populations of Gaussian functional data with same center and covariance as the
original dataset (the covariance is robustly estimated with `robustbase::covOGK`)
of size `trials_size`, and the computation of `N_trials` values for `Fvalue` 
such that the desired proportion `TPR` of observations is flagged as outliers.
The optimal value of `Fvalue` for the original population is then found as the 
average of the previously computed values `Fvalue`.
The computation of the optimal `Fvalue` at each iteration of the procedure is
carried out exploiting the zero-finding algorithm in `stats::uniroot` (Brent's
method).

The parameters to control the adjustment procedure can be passed through the
argument `adjust`, whose default is `FALSE` and otherwise is a list with 
(some of) the fields:

* `N_trials`: the number of repetitions of the adujustment procedure based on 
the simulation of a gaussisan population of functional data, each one producing
an adjusted value of F, which will lead to the averaged adjusted value `Fvalue`. 
Default is 20;
* `trial_size`: the number of elements in the gaussian population of functional 
data that will be simulated at each repetition of the adjustment procedure. 
Default is `5 * Data$N`;
* `TPR`: the True Positive Rate of outliers, i.e. the proportion of observations 
in a dataset without amplitude outliers that have to be considered outliers. 
Default is `2 * pnorm( 4 * qnorm( 0.25 ) )`;
* `F_min`: the minimum value of `Fvalue`, defining the left boundary for the 
optimisation problem aimed at finding, for a given dataset of simulated gaussian 
data associated to Data, the optimal value of `Fvalue`. Default is `0.5`;
* `F_max`: the maximum value of `Fvalue`, defining the right boundary for the 
optimisation problem aimed at finding, for a given dataset of simulated gaussian 
data associated to Data, the optimal value of `Fvalue`. Default is `5`;
* `tol`: the tolerance to be used in the optimisation problem aimed at finding, 
for a given dataset of simulated gaussian data associated to Data, the optimal 
value of `Fvalue`. Default is `1e-3`;
* `maxiter`: the maximum number of iterations to solve the optimisation problem 
aimed at finding, for a given dataset of simulated gaussian data associated to 
`Data`, the optimal value of `Fvalue`. Default is `100`;
* `VERBOSE`: a parameter controlling the verbosity of the adjustment process;


```{r fbplot_fData_adjust, eval = TRUE, collapse = TRUE, fig.align = 'center', fig.width = 7, fig.height = 4, cache = TRUE}
  fbplot( fD, adjust = list( N_trials = 10, trials_size = N, TPR = 0.007, F_min = 1, F_max = 10 ), xlab = 'grid', ylab = 'values', main = 'Adjusted functional boxplot' )
```

### Outliergram 



## References

